---
title: "Homework 3"
author: Yike Zhao
output: github_document
---

This is my solution to HW3.

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)
library(ggridges)
library(patchwork)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

### Problem 1

```{r}
data("instacart")
```
This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns.
Observations are the level of items in orders by user. There are user / order variables -- user ID, order ID, order day, and order hour. There are also item variables -- name, aisle, department, and some numeric codes. 

How many aisles, and which are most items from?

```{r}
instacart %>%
  count(aisle) %>%
  arrange(desc(n))
```

Make a plot that shows the number of items ordered in each aisle  with more than 10000 items ordered.

```{r}
instacart %>%
  count(aisle) %>%
  filter(n > 10000) %>%
  mutate(
    aisle = factor(aisle),
    aisle = fct_reorder(aisle,n)
  ) %>%
  ggplot(aes(x = aisle, y = n)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”.

```{r}
instacart %>%
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>%
  count(product_name) %>%
  mutate(rank = min_rank(desc(n))) %>%
  filter(rank < 4) %>%
  arrange(aisle, rank) %>%
  knitr::kable()
```

Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week.

```{r}
instacart %>%
  filter(product_name %in% c("Pink lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  )
```

### Problem 2

Read and tidy the accelerometers data.

```{r}
accel_df = 
  read_csv(
   "./Data/accel_data.csv"
  ) %>%
  janitor::clean_names() %>%
  mutate(
    activity_1 = as.double(activity_1),
    day = as.factor(day)
    ) %>% 
  
  pivot_longer(
    activity_1:activity_1440,
    names_to = "minute",
    names_prefix = "activity_",
    values_to = "count"
  ) %>% 

   mutate(
     day_type = case_when(
        day %in% c("Monday","Tuesday","Wednesday","Thursday","Friday") ~ "weekday",
        day %in% c("Saturday","Sunday") ~ "weekend"
    ),
    minute = as.integer(minute)
   )
```

This dataset collects the five weeks of accelerometer data of a 63 year-old male with BMI 25. The dataset contains variables: `r names(accel_df)`. It includes `r ncol(accel_df)` columns and `r nrow(accel_df)` rows.

Aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals.

```{r}
accel_df %>%
  group_by(week, day) %>% 
  summarize(day_total = sum(count))
```

There isn't any obvious trend according to the table, yet we can see that the trends for Tuesdays and Wednesdays are relatively more stable. 

Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week.
```{r}
ggplot(accel_df, aes(x = minute, y = count, color = day)) + 
  geom_smooth(se = FALSE)
```

According to the plot, we can see that there is a peak in the plot of Sunday and Friday respectively, while the peak in Sunday appears at the early half of the day, the peak in Friday appears at the end of the day. For other days in the week, the activity counts are relatively statble over the day. 

### Problem 3

Read and tidy the NY NOAA data.

```{r}
data("ny_noaa")
```

```{r}
nynoaa_df =  
  ny_noaa %>% 
  separate(date, c("year", "month", "day")) %>%
  
mutate(
  tmax = as.double(tmax)/10,
  tmin = as.double(tmin)/10,
  prcp = prcp/10
)
```

This dataset contains weather data of New York collected by the National Oceanic and Atmospheric Association (NOAA).It contains variables: `r names(nynoaa_df)`. It includes `r ncol(nynoaa_df)` columns and `r nrow(nynoaa_df)` rows.

```{r}
nynoaa_df %>% 
  drop_na(snow) %>% 
  count(snow) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4)
```

The most commonly observed snowfall values are 0, 13 and 25 mm

Make a two-panel plot showing the average max temperature in January and in July in each station across years.

```{r}
nynoaa_df %>% 
  filter(month %in% c("01","07")) %>% 
  group_by(id,month,year) %>% 
  summarize(tmax_mean = mean(tmax, na.rm = TRUE)) %>% 
  ggplot(aes(x = year, y = tmax_mean, group = id, color = month)) +
  geom_point() +
  facet_grid(. ~ month) +
  labs(
    title = "Mean Max Daily Temperature in January and July",
    x = "Year",
    y = "Mean Max Daily Temperature",
    caption ="Data from the noaa package"
  ) +
  scale_x_discrete(
    breaks = c(1980, 1985, 1990, 1995, 2000, 2005, 2010), 
    labels = c("1980", "1985", "1990", "1995", "2000", "2005", "2010"))
```

According to the graph, we can see the both of the maximum temperature of January and July fluctuated through years with the similar pattern for about every 5 years. 
There are outliers in January between 1980 and 1985, in January at around 2005, in January between 1990 and 1995, in July between 1985 and 1990.

Make a two-panel plot showing tmax vs tmin for the full dataset and the distribution of snowfall values greater than 0 and less than 100 separately by year.

```{r}
plot1_df =
nynoaa_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_bin2d() +
  theme(legend.position = "right") +
    labs(
    title = "Max Daily Temperature vs. Min Daily Temperature",
    x = "Min Daily Temperature",
    y = "Max Daily Temperature",
    caption = "Data from the noaa package"
  )
plot2_df =
nynoaa_df %>% 
  mutate(year = as.character(year)) %>% 
  filter(snow > 0 & snow < 100) %>% 
  ggplot(aes(x = year,y = snow)) + 
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5,hjust = 1))+
    labs(
    title = "Snowfall Distribution",
    x = "Time",
    y = "Snowfall",
    caption = "Data from the noaa package
    Only snowfall greater than 0 and less than 100 is included"
  )
plot1_df/plot2_df
```
